*Loading and Preprocessing the Data*
--------------------------------------
The data for the project was provided in a .zip file. The first step was to extract the file activity.csv.

The data was then read into an R data frame using the following code.

```{r}
df <- read.csv("activity.csv")
```


The total number of steps per day was calculated using the following code
```{r message=FALSE, results='hide'}
library(tidyverse)
```

```{r}
steps_per_day <- df %>% group_by(date) %>%
    summarize(Steps = sum(steps))
print(steps_per_day,n = nrow(steps_per_day))
```

Below is a histogram of the steps per day showing the frequency of the number of steps. This code was used to create the histogram.

```{r warning=FALSE}
steps_per_day %>% ggplot(aes(Steps)) +
    geom_histogram(binwidth = 1000, col = "Black", fill = "Orange") +
    ggtitle("Histogram of Steps per Day") + ylab("Frequency") +
    theme(plot.title = element_text(hjust = 0.5))
```

The mean steps per day and the median steps per day were calculated using the following code.

```{r}
mean_steps_per_day <- steps_per_day %>% summarize(mean = mean(Steps, na.rm = TRUE))
mean_steps_per_day

median_steps_per_day <- steps_per_day %>% summarize(median = median(Steps,na.rm = TRUE))
median_steps_per_day
```

