### Activity Monitoring Project ###

**Loading and Preprocessing the Data**

The data for the project was provided in a .zip file. The first step was to extract the file activity.csv.

The data was then read into an R data frame using the following code.

```{r}
df <- read.csv("activity.csv")
```

**What is mean total number of steps taken per day?**  

The total number of steps per day was calculated using the following code.
```{r message=FALSE, results='hide'}
library(tidyverse)
```

```{r}
steps_per_day <- df %>% group_by(date) %>%
    summarize(Steps = sum(steps))
print(steps_per_day,n = nrow(steps_per_day))
```

Below is a histogram of the steps per day showing the frequency of the number of steps. This code was used to create the histogram.

```{r warning=FALSE}
steps_per_day %>% ggplot(aes(Steps)) +
    geom_histogram(binwidth = 1000, col = "Black", fill = "Orange") +
    ggtitle("Histogram of Steps per Day") + ylab("Frequency") +
    theme(plot.title = element_text(hjust = 0.5))
```

The mean steps per day and the median steps per day were calculated using the following code.

```{r}
mean_steps_per_day <- steps_per_day %>% summarize(mean = mean(Steps, na.rm = TRUE))
mean_steps_per_day

median_steps_per_day <- steps_per_day %>% summarize(median = median(Steps,na.rm = TRUE))
median_steps_per_day
```

**What is the average daily activity pattern?**

The following time series plot shows the average number of steps taken per 5 minute interval.

```{r}
avg <- df %>% group_by(interval) %>% summarize(Mean = mean(steps, na.rm = TRUE)) 
avg %>% ggplot(aes(interval, Mean)) +
    geom_line(col = "Purple") +
    xlab("Time of Day (24hr time)") +
    ylab("Average Steps") +
    ggtitle("Average Number of Steps per Time of Day") +
    theme(plot.title = element_text(hjust = 0.5))
```

The maximum average number of steps was determined using the following code.

```{r}
max(avg$Mean)
```

The interval with the maximum number of steps was determined using the code below.

```{r}
avg$interval[which(avg$Mean == max(avg$Mean))]
```

**Imputing missing values**  

The total number of rows with missing steps data is calculated using the following code.

```{r}
sum(is.na(df$steps))
```

The following code replaces the NAs in the datawith the average of the appropriate 5 minute interval.

```{r}
df1 <- transform(df, steps = ifelse(is.na(steps), avg$Mean,steps))
```

The following is a histogram of steps per day

```{r}
steps_per_day <- df1 %>% group_by(date) %>%
    summarize(Steps = sum(steps))
steps_per_day %>% ggplot(aes(Steps)) +
    geom_histogram(binwidth = 1000, col = "Black", fill = "Orange") +
    ggtitle("Histogram of Steps per Day") + ylab("Frequency") +
    theme(plot.title = element_text(hjust = 0.5))
```

The mean steps per day and the median steps per day, with imputed NAs, were calculated using the following code.

```{r}
mean_steps_per_day <- steps_per_day %>% summarize(mean = mean(Steps, na.rm = TRUE))
mean_steps_per_day

median_steps_per_day <- steps_per_day %>% summarize(median = median(Steps,na.rm = TRUE))
median_steps_per_day
```

Based on these numbers, imputing the NA values did not impact the mean or median values very much.

**Are there differences in activity patterns between weekdays and weekends?**  

A new factor variable was added to the data frame characterizing the day as a weekday or the weekend.

```{r}
df1$date <- as.Date(df1$date)
df1$day <- ifelse(weekdays(df1$date) == "Saturday" | weekdays(df1$date) == "Sunday", "weekend", "weekday")
```
A panel plot was created showing the average steps per time of day for weekdays and weekends using the following code.

```{r}
avg1 <- df1 %>% group_by(interval, day) %>% summarize(Mean = mean(steps, na.rm = TRUE)) 
avg1 %>% ggplot(aes(interval, Mean)) +
    geom_line(col = "Purple") +
    facet_wrap(vars(day),nrow = 2) +
    xlab("Time of Day (24hr time)") +
    ylab("Average Steps") +
    ggtitle("Average Number of Steps per Time of Day") +
    theme(plot.title = element_text(hjust = 0.5))
```


